---
title: "Science Seekers: Basic Statistics in Psychology Module <br> Script Used to Generate the Data"
author: Kai Xiang Lim
date: 19 April, 2021
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---
``` {r setup, echo=F}
knitr::opts_chunk$set(collapse=T)
```
# Introduction to Basic Statistics in the Psychology module
This R script aims to generate a dataset for the purpose of 100SoM-IMEX Science Seeker programme

## Description of example dataset:
- SPM Maths trials results of 3 schools (long format). 
- Among them, SPM results of 2 schools are normally distributed with the same mean and sd,the 3rd school has much higher mean. 

## Steps taken
### 1. Housekeeping
Firstly we clear the workspace and load the libraries needed. 
```{r housekeeping }
rm(list=ls()) # clear the workspace

#set.seed for reproducible "random" numbers 
set.seed(1234)

#load neccesary packages
require(ggplot2)
require(psych)
require(truncnorm)
require(patchwork)
```

### 2. Data used on the slides
The data below were "generated" as illustrative purpose for the slides for this module

``` {r data used on slides}
central_tendency_eg<-c(10, 10, 10, 20, 30, 20, 10, 30, 20, 30)

mean(central_tendency_eg)

median(central_tendency_eg)

#function to get mode
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

getmode(central_tendency_eg)

(central_tendency_eg)

```

### 3. Generate data for the three schools
```{r data for schools}
#Generate random ID numbers
IdenNum<-c(100:999)
#generate a column for School A and School B
School=c(rep("SMK Ravenclaw",300),rep("SMK Slytherin",300), rep("SMK Hufflepuff", 300)) # schools as caegorical data
Sex=c(rep(c("Male","Female"),450)) # sex as binary/categorical data
data<-as.data.frame(cbind(IdenNum,School,Sex))
head(data)

#generate normally distributed scores for students from SMK Slytherin and SMK Ravenclaw separately. 

Ravenclaw_score<-round(rnorm(100, mean = 30, sd = 10),0)
Slytherin_score<-round(rnorm(100,mean = 30, sd = 10),0)
Hufflepuff_score<-Slytherin_score+40


describe(Ravenclaw_score) #min=7, max=55, mean=28
describe(Slytherin_score) #min=1, max=60, mean-30
describe(Hufflepuff_score) # min=41, max=100, mean=70
t.test(Ravenclaw_score,Slytherin_score) #Slytherin has higher mean score with p-value >0.05
t.test(Slytherin_score,Hufflepuff_score) # Hufflepuff has higher mean score with p-value <0.05
data$Score<-ifelse(data$School=="SMK Ravenclaw", Ravenclaw_score,
                   ifelse(data$School=="SMK Slytherin",Slytherin_score,
                          ifelse(data$School=="SMK Hufflepuff", Hufflepuff_score,NA)))

data$Grade<-as.factor(ifelse(data$Score<40, "G",
                             ifelse(data$Score>=40&data$Score<45, "E",
                                    ifelse(data$Score>=45&data$Score<50, "D",
                                           ifelse(data$Score>=50&data$Score<55, "C",
                                                  ifelse(data$Score>=55&data$Score<60, "C+",
                                                         ifelse(data$Score>=60&data$Score<65, "B",
                                                                ifelse(data$Score>=65&data$Score<70, "B+",
                                                                       ifelse(data$Score>=70&data$Score<80,"A-",
                                                                              ifelse(data$Score>=80&data$Score<90,"A",
                                                                                     ifelse(data$Score>=90&data$Score<=100,"A+",NA)))))))))))
```

### 4. Plotting the figures 
``` {r not-sure, echo=F}
#hist(data[data$School=="SMK Ravenclaw",]$Score)
```


##### Histogram for SMK Ravenclaw
``` {r histogram for SMK Ravenclaw}
ggplot(data,aes(x=Score))+
  geom_histogram(data=subset(data,School=="SMK Ravenclaw"), fill= "red", breaks=seq(0, 100, by=10), 
                 col="black", 
                 alpha = .2)+
  labs(title="Distribution of scores of SMK Ravenclaw students")+
  theme(plot.title = element_text(hjust = 0.5))
```


#### Histogram for SMK Slytherin
``` {r histogram for SMK Slytherin}
ggplot(data,aes(x=Score))+
  geom_histogram(data=subset(data,School=="SMK Slytherin"), fill= "blue", breaks=seq(0, 100, by=10), 
                 col="black", 
                 alpha = .2)+
  labs(title="Distribution of scores of SMK Slytherin students")+
  theme(plot.title = element_text(hjust = 0.5))
```


#### Histogram for SMK Hufflepuff
``` {r histogram Hufflepuff}
ggplot(data,aes(x=Score))+
  geom_histogram(data=subset(data,School=="SMK Hufflepuff"), fill= "green", breaks=seq(0, 100, by=10), 
                 col="black", 
                 alpha = .2)+
  labs(title="Distribution of scores of SMK Hufflepuff students")+
  theme(plot.title = element_text(hjust = 0.5))
```

#### Plotting histograms of both Slytherin and Ravenclaw together:
``` {r 2 schools histogram}
ggplot(data,aes(x=Score, fill=School)) + 
  geom_histogram(data=subset(data,School == 'SMK Ravenclaw'),col="black", breaks=seq(0, 100, by=10),alpha = 0.5) +
  geom_histogram(data=subset(data,School == 'SMK Slytherin'),col="black",breaks=seq(0, 100, by=10), alpha = 0.5)+
  labs(title="Distribution of scores of SMK Ravenclaw and SMK Slytherin")+
  theme(plot.title = element_text(hjust = 0.5))
```

#### Plotting Hufflepuff against Ravenclaw

```{r Hufflepuff vs Ravenclaw}
ggplot(data,aes(x=Score,fill=School)) +   
  geom_histogram(data=subset(data,School == 'SMK Ravenclaw'),col="black", breaks=seq(0, 100, by=10),alpha = 0.5) +
  geom_histogram(data=subset(data,School == 'SMK Hufflepuff'),col="black",breaks=seq(0, 100, by=10), alpha = 0.5)+
  labs(title="Distribution of scores of SMK Ravenclaw and SMK Hufflepuff")+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values=c("SMK Ravenclaw"="Red","SMK Hufflepuff"="Green" ))

```

#### Boxplot for SMK Hufflepuff
``` {r boxplotHufflepuff}

boxplot(subset(data, School=="SMK Hufflepuff")$Score,main="Boxplot: SMK Hufflepuff Scores", col="green")
```


#### Boxplot for all schools
``` {r boxplot all schools}
ggplot(data,aes(x=Score,y=School, fill=School))+
  geom_boxplot()+ylab("")+labs(title="Boxplot for all schools")
```

### 5. Identifying data type
#### Recode data such that grade is ordinal variable, score is continuous variable, sex is categorical variable

Reorder grade and plot bar chart
``` {r recode data}
data$Grade <- ordered(data$Grade, levels = c("G", "E", "D", 
                                             "C", "C+", "B",
                                             "B+", "A-", "A",
                                             "A+"))
plot(data$Grade, col=c("black", "red","orange","pink", "yellow","green",
                       "blue", "violet", "purple", "brown"), 
     main="Distribution of grades of all three schools")
```

### 6. Mean, mode, median 
#### Showing mean score for each school 
``` {r mean scores}
table<-aggregate(data$Score, list(data$School), mean)
colnames(table)<-c("School", "Mean")
table

table2<-aggregate(data$Score, list(data$School), median)
colnames(table2)<-c("School", "Median")
table2

merge(table,table2, by="School")

```


#### Get mode
``` {r get mode}
getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}

table3<-aggregate(as.factor(data$Grade), list(data$School), getmode)
colnames(table3)<-c("School", "Mode")
table3
```


### 6. Save the data
```{r save data, eval=FALSE, include=T}
setwd("/Users/kai/OneDrive - King's College London/PhD/outside_PhD/iMEX_SoM")
xlsx::write.xlsx(data, "Maths_trial_data.xlsx")

```

